Think deep for reef resilience

### 2025-08-07 paper/paper_draft.qmd
Shift description of depth estimation to earlier section
@eco-ben Could you double check the mentioned number of reefs for consistency in the text for depth estimation please?

Shifting text earlier up

## Coral reef ecosystem model - ADRIA-CoralBlox

@@

### Depth estimation

@@

Overall, these findings support the hypothesis that reefs at greater depths will experience
refugia benefits and more prolonged resilience under climate change compared to shallower
reefs. Additionally, Lui, Callaghanand Baldock [-@liu2024] suggest that rubble instability
is inversely correlated with reef depth, possibly limiting the juvenile recruitment capacity
of shallow reefs. A caveat to these depth-resilience findings is the quality of the
bathymetry data used to calculate median reef depth. These data (GBRMPA [-@gbr10gb2020]) are
derived from satellite imagery and are limited in their alignment with the established
GBRMPA reef outlines used in the investigation to characterise reef locations. As a result,
842 reefs (22% of 3806 reefs) were removed from the depth-related analyses due to having
less than 5% coverage of depth data. Additionally, deriving estimates of depth from
satellite imagery is prone to high degrees of uncertainty. Data adopted for the given
assessment are said to be accurate up to 15m from Mean Sea Level [@gbr10gb2020].
Furthermore, an estimated median depth is determined for each reef, using the established
GBRMPA reef outlines to provide an indication of “true” reef extent. Therefore, the
estimates of reef depths are taken as a useful indication rather than an accurate and
precise report of reef depths across the GBR.

>>>>

## Depth estimation

\[Below moved from discussion - to be reworked for flow\]

These data (GBRMPA [-@gbr10gb2020]) are derived from satellite imagery and can be limited in
their alignment with the established GBRMPA reef outlines used in the investigation to
characterise reef locations. As a result, 842 reefs (22% of 3806 reefs) were removed from
the depth-related analyses due to having less than 5% coverage of depth data. **\[TI to BG:
is this number of reefs consistent?\]**

Additionally, deriving estimates of depth from satellite imagery is prone to high degrees of
uncertainty. Data adopted for the given assessment are said to be accurate up to 15m from
Mean Sea Level [@gbr10gb2020]. Furthermore, an estimated median depth is determined for each
reef, using the established GBRMPA reef outlines to provide an indication of “true” reef
extent. Therefore, the estimates of reef depths are taken as a useful indication rather than
an accurate and precise report of reef depths across the GBR.

## ADRIA-CoralBlox modelling

@@

@@

These findings support the hypothesis that reefs at greater depths will exhibit
characteristics of refugia and more prolonged structural resilience under climate change
compared to shallower reefs. Additionally, Lui, Callaghan and Baldock [-@liu2024] suggest
that rubble instability is inversely correlated with reef depth, possibly limiting the
juvenile recruitment capacity of shallow reefs. A caveat to these depth-resilience findings
is the quality of the bathymetry data used to calculate median reef depth. As detailed
previously, estimates of median reef depth are derived from satellite imagery (at $10m^2$
resolution) which can be subject to high degrees of uncertainty, both in terms of the
estimated depth at for each pixel as well as their alignment to the established GBRMPA reef
outlines.
